---
title: "How H2O works in R"
author: "Manish Barnwal"
date: "December 1, 2015"
output: html_document
---


# Big Data Science with H2O in R

## H2O
H2O is "The Open Source In-Memory, Prediction Engine for Big Data Science"

## Install the R package

## The following two commands remove any previously installed H2O packages for R.
```{r}
# if ("package:h2o" %in% search()) { detach("package:h2o", unload=TRUE) }
# if ("h2o" %in% rownames(installed.packages())) { remove.packages("h2o") }
#  
# install.packages("h2o")
```

## Next, we launch h2o from R
```{r}
# library(h2o)
# localH2o <- h2o.init()

```
## The above will initialize your h2o cluster. Now there are other parameter in 'h2o.init' but for now let's go with the default settings. Few important parameters to 'h2o.init()' are
1. 'nthreads': Number of cores of the computer you want to use.

2. 'max_mem_size': The total RAM size allocated to the cluster. 

## Check if connection is working
```{r}
# h2o.clusterInfo(localH2o)
```

## Interaction of R with the cluster
1. The data is not saved in the R workspace. All data munging occurs in h2o instance

2. When the user makes a request, R queries the server via the REST API, which returns a JSON file with the relevant information that R then displays in the console


## Machine learning algorithms currently available in h2o
1. GLM 

2. GBM

3. Kmeans 

4. Naive Bayes

5. Principal Components Analysis

6. Deep Learning

7. Random Forests

## Demonstration on a dataset
### Classification using Random forest 
Refer 'DhackCodeDemo.R'. All you need to do is change your working directory and put the data files into that directory. The code will run fine after this.

## Interaction between UI and Rstudio

1. You can load the objects that you have created in your UI mode to your Rstudio using

### get model from the UI
```{r}
# rf.model = h2o.getModel("drf-e4326760-bb35-49e8-a7ae-3cd6d572ec29")
 
```

Now 'rf.model' is your model and you can use this to predict on your test dataset. Note that 'rf.model' will work only till your cluster is on. The moment you shutdown your cluster, this model will be lost. 
There should be a way out for this problem or else everytime you will have to run your model from scratch or the other way out is to not shutdown your cluster at all. Not feasible.

2. There is a way to save your model to your disk. And you can use this model even when your cluster is shutdown. 

### Saving model to your local disk. 
```{r}
# h2o.saveModel(rf.model, dir = "file:///C:/Users/mbarnwa/Desktop", force = T) 
```

### Once you have saved your model. You can load the model and use it to predict
```{r}
# loadedModel <- h2o.loadModel("file:///C:/Users/mbarnwa/Desktop/randomForestExample")

```

**Note the extra 'file:///; in 'h2o.saveModel()' and 'h2o.loadModel()'** 

### That is all for today!

